---
title: "Association_Rules"
author: "Nathan_Njonge"
date: "7/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PROBLEM DEFINITION

You are a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). Your project has been divided into four parts where you'll explore a recent marketing dataset by performing various unsupervised learning techniques and later providing recommendations based on your insights.

```{r}
# load our dataset
df <- read.csv("http://bit.ly/SupermarketDatasetII")
```

```{r}
# preview the top of our dataset
head(df)
```

```{r}
# preview the bottom of our dataset
tail(df)
```

```{r}
# view the dimensions of our dataset
dim(df)
```

> Our dataset has 7500 rows and 20 columns.

```{r}
# summary of our dataset
summary(df)
```
```{r}
# check for missing values
colSums(is.na(df))
```

> There appears to be missing values in the olive.oil column

```{r}
# omit missing values
df2 <- subset(df, select=-c(olive.oil))
colSums(is.na(df2))
```
```{r}
# find any duplicated rows in our dataset
duplicated_rows <- df2[duplicated(df2),]
head(duplicated_rows)
```

> There is presence of duplicated rows in our dataset

```{r}
# dropping duplicated items
unique_items <- df2[!duplicated(df2), ]
# view the unique items
head(unique_items)
```

## Implementing the Solution

```{r}
# loading the libraries
library(arules)
library(arulesViz)
library(tidyverse)
library(readxl)
library(knitr)
library(ggplot2)
library(lubridate)
library(plyr)
library(dplyr)
```


```{r}
# store this unique items into a csv
write.csv(unique_items,"D:/IP_WEEK_14/analysis.csv", quote = FALSE, row.names = FALSE)
```


```{r}
# convert our unique items into an object of the transaction class
tr <- read.transactions("D:/IP_WEEK_14/analysis.csv", format = 'basket', sep=',')
```

```{r}
# View the transaction object
tr
```

```{r}
# summary of the transaction object
summary(tr)
```


```{r}
# Create an item frequency plot for the top 20 items
if (!require("RColorBrewer")) {
# install color package of R
install.packages("RColorBrewer")
#include library RColorBrewer
library(RColorBrewer)
}
itemFrequencyPlot(tr,topN=20,type="absolute",col=brewer.pal(8,'Pastel2'), main="Absolute Item Frequency Plot")
```
> Our absolute item frequency plot gives us the numeric frequencies of each item independently where mineral water appears to have had the highest frequency. 

```{r}
# item frequency plot for the top 20 items using relative type
itemFrequencyPlot(tr,topN=20,type="relative",col=brewer.pal(8,'Pastel2'),main="Relative Item Frequency Plot")
```
> Our relative item frequency plot tells us how many times these items have appeared as compared to others, whereby mineral water has appeared more compared to the others. 

## Generating Rules

```{r}
# Min Support as 0.001, confidence as 0.8.
association.rules <- apriori(tr, parameter = list(supp=0.001, conf=0.8,maxlen=10))
```


```{r}
inspect(association.rules[1:10])
```

> Using the above output we can be able to make our analysis. For example 100% of the customers who bought 'nonfat milk,turkey' also bought 'mineral water'.

## Limiting the number and size of rules

```{r}
shorter.association.rules <- apriori(tr, parameter = list(supp=0.001, conf=0.8,maxlen=3))
```

## Removing redundant rules

```{r}
# remove rules that are subsets of larger rules

# get subset rules in vector
subset.rules <- which(colSums(is.subset(association.rules, association.rules)) > 1) 

length(subset.rules)
```

```{r}
# remove subset rules.
subset.association.rules. <- association.rules[-subset.rules]
```


# Challenging the solution

## Finding rules related to chocolate

```{r}
# find what customers buy before buying 'chocolate'
chocolate.association.rules <- apriori(tr, parameter = list(supp=0.001, conf=0.8),appearance = list(default="lhs",rhs="chocolate"))
```


```{r}
inspect(head(chocolate.association.rules))
```

> Above is the output of what people buy before buying chocolate.




